---
title: "Ashbury - Census results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(readxl)
library(data.table)
library(ggplot2)
#download.file("http://www.censusdata.abs.gov.au/CensusOutput/copsub2016.NSF/All%20docs%20by%20catNo/2016~Community%20Profile~120031392/$File/TSP_120031392.zip","cna.zip")

#unzip("cna.zip")
file <- "TSP_120031392.XLS"
ages <- setDT(read_excel(file, sheet = "T 01", range = "A13:L24"))
long_age <- melt(ages,id=1,measure.vars = c(2:4,6:8,10:12))
long_age[, year := rep(c(2006,2011,2016),each=3*length(unique(`Age group(a):`)))]
long_age[, sex := rep(c("male","female","persons"),each=length(unique(`Age group(a):`)))]
long_age[, Age := gsub(" years","", `Age group(a):`)][, Age:= gsub(" and over"," +",Age)]
long_age[, AgeFactor := factor(Age, levels = unique(Age))]
```

# Canterbury North / Ashbury SA2

Source: Community profile: http://www.censusdata.abs.gov.au/census_services/getproduct/census/2016/communityprofile/120031392?opendocument 


## Population by Age for Census years
```{r}

ggplot(long_age[sex=='persons'], aes(x= year, y = value,colour=AgeFactor)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2006,2011,2016)) +
  scale_color_brewer(palette = "Spectral") +
  labs(x = "Census year", y = "Population", colour = 'Age') +
  theme_bw()
```

## Population by Age and Sex
```{r}
ggplot(long_age[sex!='persons'], aes(x = AgeFactor, y = value)) +
  geom_bar(stat="identity") +
  facet_grid(year ~ sex) +
  coord_flip() +
  labs(x = "Age", y = "Population") +
  theme_bw()
```

#Population Growth
```{r}
age_growth <- long_age[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(AgeFactor,sex)]
```

Note declines from 2006 - 2011 but large growth in 25-34 and and similar growth in  by 5-14, 35-54 age group.

```{r,results='asis'}
ggplot(age_growth[sex=='persons'], aes(x = AgeFactor, y = growth)) + 
  geom_bar(stat="identity",aes(fill = factor(growth>0))) +
  facet_wrap(~year) + 
  coord_flip() +
  scale_y_continuous(label = scales::percent) +
  labs(y = "% Growth in from previous Census",
       x = "Age group") +
  theme_bw() +
  scale_fill_brewer(palette ="Set1") +
  theme(legend.position="none")


knitr::kable(age_growth[sex=='persons'][year==2016][,list("Age group" = AgeFactor, Percent = scales::percent(growth))])
```

## Population growth by gender

Note losses 2006-2011 in 15-44 women. Now growing (2011-2016)

```{r}
ggplot(age_growth[sex!='persons'], aes(x = AgeFactor, y = growth)) + 
  geom_bar(stat="identity",aes(fill = factor(growth>0))) +
  facet_grid(sex~year) + 
  coord_flip() +
  scale_y_continuous(label = scales::percent) +
  labs(y = "% Growth in from previous Census",
       x = "Age group") +
  theme_bw() +
  scale_fill_brewer(palette ="Set1") +
  theme(legend.position="none") 
```

# Families and family structure

```{r}
# Marital status and children

families = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 27", range = "B17:F18",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 27", range = "B27:F28",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 27", range = "B37:F38",col_names = FALSE)),
  id = "year"
)

long_families <- melt(families, id=1)[,
  label := rep(c(1,2,3,"4+","Total"), each=6)][,
  label2 := rep_len(c("Two","One"),.N)]

total_families <- long_families[,list(N = sum(value)),by=list(label,year)]
family_growth <- total_families [,  list(year = tail(year, -1), growth = diff(N) / head(N,-1)), by = list(label)]
```

## Number of families by children structure  2016

```{r }
ggplot(total_families[year==2016][label !='Total'], aes(x = label, y= N,fill=label)) + 
  geom_bar(stat='identity')+
  coord_flip() +
  scale_fill_brewer(palette="Set1") +
  labs(x = "Children", y= "Families") +
  theme_bw() + theme(legend.position = "none")
```

## Growth in families

Note 7.7% more families in 2016, particularly 1-2 children families (i.e. "young" /"smaller" / "beginning")

```{r results = 'asis'}
ggplot(family_growth,aes(x = label, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity") + 
  facet_wrap(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census","Number of Children") +
  theme_bw() + theme(legend.position = 'none')

knitr::kable(family_growth[year==2016][,list(Children  = label, growth = scales::percent(growth))])

```


# English proficiency
Note growth in low-English proficiency in 15-34 year olds
```{r}
english = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 11", range = "A13:I21",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 11", range = "A27:I35",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 11", range = "A41:I49",col_names = FALSE)),
  id = "year"
)[,!'X__3']
english_long <- melt(english, id =1:2)

labels <- c("English_only","Very well","Not very well","Not stated proficiency","Total non english","Not stated Language","Total")
english_long[, label := variable]
levels(english_long[['label']]) <- labels
english_long[,Age := gsub(" and over", " +",gsub(X__1,pattern = " years",replacement = ""))]
english_long[,AgeFactor := factor(Age, levels = unique(Age))]

not_well <- english_long[label=="Not very well"]
growth_not_well <- not_well[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(AgeFactor)]
ggplot(not_well,aes(x = AgeFactor, y= value)) + 
  geom_bar(stat='identity') + 
  facet_wrap(~year) +coord_flip() +
  theme_bw() +
  labs(x = "Age", y= "Number of people who do not speak English well or at all")


ggplot(growth_not_well,aes(x = AgeFactor, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity") + 
  facet_wrap(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Age",title = "People who do not speak English well or at all") +
  theme_bw() + theme(legend.position = 'none')
```


## Religion

```{r}

religion = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 12a", range = "A13:I45",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 12b", range = "A13:I45",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 12c", range = "A13:I45",col_names = FALSE)),
  id = "year"
)

religion[, Group := cumsum(X__1 %in% c("Buddhism","Christianity:","Hinduism","Judaism","Other Religions:","Secular Beliefs and Other Spiritual Beliefs","Religious affiliation not stated(d)")),by=year]
GroupNames <- c("Buddhism","Christianity","Hinduism","Judaism","Other","No religion", "Not stated")
religion_long <- melt(religion, id = c(1,2,11))
total_religion <- religion_long[X__1 !='Total', list(value = sum(value,na.rm=TRUE)),by=list(year, Group= factor(Group,labels=GroupNames))]

religion_by_age <- religion_long[X__1 !='Total', list(value = sum(value,na.rm=TRUE)),by=list(year, Group,variable)]
religion_by_age[, Religion := factor(Group, labels = GroupNames)]
religion_by_age[, AgeFactor := variable]
levels(religion_by_age[['AgeFactor']]) <- growth_not_well[,levels(AgeFactor)]
ggplot(total_religion, aes(x = Group, y= value,fill=year)) + 
  geom_bar(stat='identity',position="dodge") + 
  coord_flip() +
  theme_bw() +
  labs(x = "Religion", y= "Number") +
  scale_fill_brewer(palette="Set1")

growth_religion  <- total_religion[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(Group)]
growth_religion_age  <- religion_by_age[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(Religion,AgeFactor)]
ggplot(growth_religion,aes(x = Group, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity") + 
  facet_wrap(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Religion") +
  theme_bw() + theme(legend.position = 'none')
```

## Religion by age



```{r}
ggplot(religion_by_age, aes(x = Religion, y= value,fill=AgeFactor)) + 
    geom_bar(stat='identity',position="dodge") + 
    coord_flip() + facet_wrap(~year) +
    theme_bw() +
    labs(x = "Religion", y= "Number") +
    scale_fill_brewer(palette="Set1","Age group") 


ggplot(growth_religion_age,aes(x = AgeFactor, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity",position="dodge") + 
  facet_grid(Religion~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Age") +
  theme_bw() + theme(legend.position = 'none')

```


## Christian Demoniations:
Within Christians: 
Note amongst wide spread "reduction" 2011-2016 small growth for 15-34 Anglicans.
(Also, Baptist doing well from a low base)

```{r}
Christians <- religion_long[X__1 !='Total'&Group==2]
Christians[, Denomination := X__1][!X__1 %in%c("Anglican","Baptist","Eastern Orthodox","Catholic"), Denomination := "Other"]

by_denom_age <- Christians[,list(value = sum(value, na.rm =TRUE)),by=list(year,Denomination,AgeFactor = variable)]
levels(by_denom_age[['AgeFactor']]) <- growth_religion_age[,levels(AgeFactor)]
by_denom <- Christians[,list(value = sum(value, na.rm =TRUE)),by=list(year,Denomination)]

growth_denom <- by_denom[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(Denomination)]
growth_denom_age  <- by_denom_age[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(Denomination,AgeFactor)]

ggplot(by_denom, aes(x = Denomination, y= value,fill=year)) + 
  geom_bar(stat='identity',position="dodge") + 
  coord_flip() +
  theme_bw() +
  labs(x = "Christian Denomination", y= "Number") +
  scale_fill_brewer(palette="Set1")

ggplot(growth_denom,aes(x = Denomination, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity") + 
  facet_wrap(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Denomination") +
  theme_bw() + theme(legend.position = 'none')


ggplot(by_denom_age, aes(x = Denomination, y= value,fill=AgeFactor)) + 
    geom_bar(stat='identity',position="dodge") + 
    coord_flip() + facet_wrap(~year) +
    theme_bw() +
    labs(x = "Denomination", y= "Number") +
    scale_fill_brewer(palette="Set1","Age group") 


ggplot(growth_denom_age,aes(x = AgeFactor, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity",position="dodge") + 
  facet_grid(Denomination~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Age") +
  theme_bw() + theme(legend.position = 'none')

```

# Income for families

```{r}

incomeCouples = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 22a", range = "A13:F28",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 22a", range = "A34:F49",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 22b", range = "A13:F28",col_names = FALSE)),
  id = "year"
)

incomeOne = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 23a", range = "A13:F28",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 23a", range = "A34:F49",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 23b", range = "A15:F30",col_names = FALSE)),
  id = "year"
)

income <- rbindlist(list(Couple  = incomeCouples, One = incomeOne), id = "Parent")
setnames(income,"X__1","Income")

income_long <- melt(income, id =c(1,2,3))
levels(income_long[['variable']]) <- c("1","2","3","4+","Total")
income_long[ ,IncomeFactor := factor(Income,levels = unique(Income))]

income_total <- income_long[variable=='Total',list(value=sum(value)),by=list(year,IncomeFactor)]

growth_income <- income_total[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(IncomeFactor)]

ggplot(income_long[variable=="Total"][Income %like% "[,/]"], aes(x = IncomeFactor, y = value,fill=Parent)) +
  geom_bar(stat='identity', position = 'stack') +
  facet_wrap(~year) +
  coord_flip() +
  scale_fill_brewer(palette='Set1') +
  labs(x = "Income bracket",y="Number") +
  theme_bw()


ggplot(growth_income[IncomeFactor%like% "[,/]"],aes(x = IncomeFactor, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity",position="dodge") + 
  facet_grid(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census",x= "Income bracket") +
  theme_bw() + theme(legend.position = 'none')