---
title: "Croydon Park / Enfield - Census results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(data.table)
library(ggplot2)
download.file("http://www.censusdata.abs.gov.au/CensusOutput/copsub2016.NSF/All%20docs%20by%20catNo/2016~Community%20Profile~120031393/$File/TSP_120031393.zip","cpe.zip")

unzip("cpe.zip")
file <- "TSP_120031393.XLS"
ages <- setDT(read_excel(file, sheet = "T 01", range = "A13:L24"))
long_age <- melt(ages,id=1,measure.vars = c(2:4,6:8,10:12))
long_age[, year := rep(c(2006,2011,2016),each=3*length(unique(`Age group(a):`)))]
long_age[, sex := rep(c("male","female","persons"),each=length(unique(`Age group(a):`)))]
long_age[, Age := gsub(" years","", `Age group(a):`)][, Age:= gsub(" and over"," +",Age)]
long_age[, AgeFactor := factor(Age, levels = unique(Age))]
```

# Croydon Park /Enfield SA2

Source: Community profile: http://www.censusdata.abs.gov.au/census_services/getproduct/census/2016/communityprofile/120031393?opendocument 


## Population by Age for Census years
```{r}

ggplot(long_age[sex=='persons'], aes(x= year, y = value,colour=AgeFactor)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2006,2011,2016)) +
  scale_color_brewer(palette = "Spectral") +
  labs(x = "Census year", y = "Population", colour = 'Age') +
  theme_bw()
```

## Population by Age and Sex
```{r}
ggplot(long_age[sex!='persons'], aes(x = AgeFactor, y = value)) +
  geom_bar(stat="identity") +
  facet_grid(year ~ sex) +
  coord_flip() +
  labs(x = "Age", y = "Population") +
  theme_bw()
```

#Population Growth
```{r}
age_growth <- long_age[,  list(year = tail(year, -1), growth = diff(value) / head(value,-1)), by = list(AgeFactor,sex)]
```

Note declines from 2006 - 2011 but large growth in 25-34 and and similar growth in  by 5-14, 35-54 age group.

```{r,results='asis'}
ggplot(age_growth[sex=='persons'], aes(x = AgeFactor, y = growth)) + 
  geom_bar(stat="identity",aes(fill = factor(growth>0))) +
  facet_wrap(~year) + 
  coord_flip() +
  scale_y_continuous(label = scales::percent) +
  labs(y = "% Growth in from previous Census",
       x = "Age group") +
  theme_bw() +
  scale_fill_brewer(palette ="Set1") +
  theme(legend.position="none")


knitr::kable(age_growth[sex=='persons'][year==2016][,list("Age group" = AgeFactor, Percent = scales::percent(growth))])
```

## Population growth by gender

Note largest growth in 85 . No other obvious changes.

```{r}
ggplot(age_growth[sex!='persons'], aes(x = AgeFactor, y = growth)) + 
  geom_bar(stat="identity",aes(fill = factor(growth>0))) +
  facet_grid(sex~year) + 
  coord_flip() +
  scale_y_continuous(label = scales::percent) +
  labs(y = "% Growth in from previous Census",
       x = "Age group") +
  theme_bw() +
  scale_fill_brewer(palette ="Set1") +
  theme(legend.position="none") 
```

# Families and family structure

```{r}
# Marital status and children

families = rbindlist(
  list(`2006` = read_excel(file, sheet = "T 27", range = "B17:F18",col_names = FALSE),
       `2011` = read_excel(file, sheet = "T 27", range = "B27:F28",col_names = FALSE),
       `2016` = read_excel(file, sheet = "T 27", range = "B37:F38",col_names = FALSE)),
  id = "year"
)

long_families <- melt(families, id=1)[,
  label := rep(c(1,2,3,"4+","Total"), each=6)][,
  label2 := rep_len(c("Two","One"),.N)]

total_families <- long_families[,list(N = sum(value)),by=list(label,year)]
family_growth <- total_families [,  list(year = tail(year, -1), growth = diff(N) / head(N,-1)), by = list(label)]
```

## Number of families by children structure  2016

```{r }
ggplot(total_families[year==2016][label !='Total'], aes(x = label, y= N,fill=label)) + 
  geom_bar(stat='identity')+
  coord_flip() +
  scale_fill_brewer(palette="Set1") +
  labs(x = "Children", y= "Families") +
  theme_bw() + theme(legend.position = "none")
```

## Growth in families

Largest growth in 2 and 4 children families (existing families having more children?)

Fewer

```{r results = 'asis'}
ggplot(family_growth,aes(x = label, y= growth, fill = growth >0))+ 
  geom_bar(stat="identity") + 
  facet_wrap(~year) +coord_flip() + 
  scale_fill_brewer(palette="Set1") +
  scale_y_continuous(label=scales::percent)+
  labs(y = "% Growth in from previous Census","Number of Children") +
  theme_bw() + theme(legend.position = 'none')

knitr::kable(family_growth[year==2016][,list(Children  = label, growth = scales::percent(growth))])

```